---
title: "Implied volatility for Apple options"
author: "Robert McDonald"
output:
  pdf_document:
    toc: yes
  html_document:
    toc: yes
  word_document: default
---

```{r, echo=FALSE}
library(knitr)
ticker <- 'Apple'
opts_chunk$set(echo=FALSE)

```

# Introduction

This is document uses Optionmetrics data to examine Apple options from
September 2013 to August 2014. The document is constructed using
RMarkdown, and processed using RStudio. The original RMarkdown
document is on Dropbox and available
[here](https://www.dropbox.com/s/9qwj4s62bacqsyl/plotallimpvol.Rmd?dl=0)

# Volume and Open Interest for `r ticker`, Sept 2013 - August 2014

The following plots show option volume and open interest for 
`r ticker`. The vertical lines in each plot mark earnings 
announcement dates.

Apple shares split 7-for-1 on June 9, 2014. In order to provide continuity, the volume and open interest amounts are normalized by dividing by 7 after that date.


```{r, results='asis'}
## file to read optionmetrics data

# setwd('C:/RDerivatives')
startdate <- as.Date('2013/09/01')
enddate <- as.Date('2014/08/30')
## fn <- 'volume.csv'
fn <- 'aapl_opts.csv'
fnstk <- 'aapl_stock.csv'
loadfile <- TRUE
## loadfile <- FALSE

earndates <- as.Date(c('2015/01/27', '2014/10/20', '2014/04/23',
	'2014/07/22', '2014/01/27', '2013/10/28', '2013/07/23',
	'2013/04/23', '2013/01/23', '2012/10/25', '2012/07/24',
	'2012/04/24', '2012/01/24'))

earndates <- sort(earndates)
splitdate <- as.Date('2014/06/09') ## 7 for 1

if (loadfile) {
    load('appldata.Rdata')
} else {
    x <- read.csv(fn, stringsAsFactors=FALSE, nrows=500)
    cc <- sapply(1:length(x), function(i) class(x[,i]))
    x <- read.csv(fn, stringsAsFactors=FALSE, colClasses=cc)
    x$strike_price <- x$strike_price/1000
    x$date <- as.Date(x$date)
    x$exdate <- as.Date(x$exdate)
    x[x$last_date=="", 'last_date'] <- NA
    x$last_date <- as.Date(x$last_date)
    x$daysexp <- as.numeric(x$exdate - x$date)
    stk <- read.csv(fnstk, stringsAsFactors=FALSE)[,c(2,3)]
    stk$date <- as.Date(stk$date,"%Y/%m/%d")
    x <- merge(x, stk[, c('date', 'close')],
                   by.x='date', by.y='date', all.x=TRUE)    
    save(x, file='appldata.Rdata')
}
x <- x[x$date <= enddate & x$date >= startdate, ]

```

```{r, results='asis'}
## plot volume and open interest over the year by puts/calls,
## moneyness, expiry
y <- with(x, aggregate(cbind(open_interest=open_interest/cfadj, 
                             volume=volume/cfadj),
               by=list(cp_flag = cp_flag, date = date), FUN=sum))
yc <- y[y$cp_flag=='C', ]
yp <- y[y$cp_flag=='P', ]
```

## Call and Put Option Volume

```{r, echo=FALSE, results='asis'}
plot(yc$date, yc$volume, cex=0.5,
     xlab='Date', main='Call Volume')
abline(v=earndates, lty=2)
plot(yp$date, yp$volume, cex=0.5,
     xlab='Date', main='Put Volume')
abline(v=earndates, lty=2)
```

## Call and Put Open Interest

```{r, echo=FALSE, results='asis'}
plot(yc$date, yc$open_interest, cex=0.5,
     xlab='Date', main='Call Open Interest')
abline(v=earndates, lty=2)
plot(yp$date, yp$open_interest, cex=0.5,
     xlab='Date', main='Put Open Interest')
abline(v=earndates, lty=2)
```


# Implied volatilities for Apple, Sept 2013 - August 2014

The following plots show implied volatilities for every day in the period. The
plots show only implied volatilities for 5 sets of options. These are
determined by sorting the existing times to expiration on each day, and
plotting implied volatilities for the 2nd nearest expiration, 4th nearest, out to the 10th nearest time to expiration. There are always current week and
following week options, so the nearest to expiration options 
have times to expiration varying from 7 to 11 days.

The dashed vertical line on each plot is that day's closing stock price.

Earnings announcement dates for `r ticker` were 
`r earndates[8:10]` and `r earndates[11]`. 
These are noted in the plot title for those days.


```{r, echo=FALSE, results='asis'}
xp <- x[x$cp_flag=='P', ]
wk <- c(2, 4, 6, 8, 10)
colors <- c('red', 'green', 'blue', 'magenta', 'black')
alldates <- sort(unique(xp$date))
previousdate <- as.POSIXlt(alldates[1])
cat('\n\n##', strftime(previousdate, '%B, %Y'), '\n\n')
for (i in alldates) {
    ip <- as.POSIXlt(as.Date(i, origin="1970-01-01"))  
    if (ip$mon != previousdate$mon) {
      cat('\n\n## ', strftime(ip, '%B, %Y'), '\n\n')
    }
    earntext <- ''
    if (as.Date(i, origin = "1970-01-01") %in% earndates) earntext <- '(Earnings Announcement)'
    tmp <- subset(xp[xp$date==i,])
    exdate <- sort(unique(tmp$exdate))
    xvar <- tmp$strike_price[tmp$exdate==exdate[wk[1]]]
    yvar <-tmp$impl_volatility[tmp$exdate==exdate[wk[1]]] 
    plot(xvar, yvar, cex=0.25, col=colors[1],
         main=paste('Apple puts: ', tmp$date[1], earntext, 
         		   ' (Price = ', tmp$close[1], ')', sep=''),
         xlab='Strike', ylab='Implied volatility',
         xlim=c(.6*min(xvar), 1.4*max(xvar)),
         ylim=c(0, 1.5))
    for (j in 2:length(wk)) {
        points(tmp$strike_price[tmp$exdate==exdate[wk[j]]],
         tmp$impl_volatility[tmp$exdate==exdate[wk[j]]], cex=0.25, col=colors[j]) 
    }
    abline(v=tmp$close[1], lty=2)
    legend('topleft', legend=c(paste(exdate[wk] - tmp$date[1], 'days to expiry')),
           col=colors, pch=1, cex=0.5)
    previousdate <- ip
}

```
